---
title: "Analysis on the relationship between Transmission Type and MPG"
output:
  pdf_document: default
  html_notebook: default
---

# Introduction

In this analysis I will try to find if there is a relationship between a set of variables and the miles per galons of a set of cars. To do this, we will use the *mtcars* dataset that is included in R.

The data set includes the following:

A data frame with 32 observations on 11 (numeric) variables.

* [, 1]	mpg	Miles/(US) gallon
* [, 2]	cyl	Number of cylinders
* [, 3]	disp	Displacement (cu.in.)
* [, 4]	hp	Gross horsepower
* [, 5]	drat	Rear axle ratio
* [, 6]	wt	Weight (1000 lbs)
* [, 7]	qsec	1/4 mile time
* [, 8]	vs	Engine (0 = V-shaped, 1 = straight)
* [, 9]	am	Transmission (0 = automatic, 1 = manual)
* [,10]	gear	Number of forward gears
* [,11]	carb	Number of carburetors

The course project asks two questions:

* Is an automatic or manual transmission better for MPG?
* Quantify the MPG difference between automatic and manual transmissions

# Exploratory Data Analysis

We begin by loading the necessary libraries and the data set: 

```{r}
library(ggplot2)
data(mtcars)
mtcars$cyl = as.factor(mtcars$cyl)
mtcars$vs = as.factor(mtcars$vs)
mtcars$am = as.factor(mtcars$am)
```

```{r}
head(mtcars)
```
```{r}
str(mtcars)
```

We can then use some graphics to find a visible relationship:
```{r}
pairs(mtcars)
```

Seeing the pairs plot, we can identify the following for mpg:

* A negative relationship with disp, hp, cyl and wt.
* A positive relationship with drat and qsec.
* A clear difference between vs and am.

Since the scope of this study only includes the transmission type, we will only analyse the am variable

```{r}
g <- ggplot(data = mtcars, aes(x = factor(am), y = mpg))
g + geom_boxplot() + ggtitle("Miles per Galon vs Transmission Type") + xlab("Transmission Type [0: auto ; 1: manual]") + ylab("Miles per Galon")
```

Using this boxplot, we could assume that there is a very clear difference between the fuel efficiency of automatic and manual cars, showing that manual transmissions lead to a better mileage. We wan't to test if there is a statistically significant difference in the gas mileage of both types of transmissions. 

# Statistical Analysis
To do this, we perform a T test to compare the medians
```{r}
t.test(mpg ~ am, data = mtcars)
```

Using a p-value of 0.025 (two-sided test) shows that there is a significant difference in the medians of both groups, meaning that manual cars do have a better gas mileage than those with automatic transmissions.

# Regression Model

To find a proper regression model, we start with a basic linear model using mpg as the outcome and am as the regression:

```{r}
fit <- lm(mpg ~ am, data = mtcars)
summary(fit)
```

We haven an adjusted R-squared of 0.3385, meaning that our model is not very accurate and that am only accounts for around 33% of the increase in mpg. We can test and overfit this model using all variables:

```{r}
fit_all <- lm(mpg ~ ., data = mtcars)
summary(fit_all)
```

Here we have a better R-squared, but using a p-value of 0.05 shows that none of the variables are significant, which is a sign of overfitting. At this point we use the *step* function to iteratively find a model that better fits our data:

```{r}
fit_step <- step(fit_all, trace = F)
summary(fit_step)
```

All variables in this model have a p-value bigger than 0.05, meaning they are statistically significant. The step function chose the variables *wt*,*qsec* and *am* for our new model.
The confidence interval for this assumption is:

```{r}
confint(fit_step)["am1",]
```
```{r}
par(mfrow = c(2,2))
plot(fit_step)
```

The Residuals vs. Fitted plot shows that the residuals are uncorrelated with the fitted values and the Normal Q-Q plot shows that the distribution is roughly normal

# Conclusions
After analyzing the data we can conclude that there is a strong relationship between transmission type and gas mileage. Other significant variables are the Weight and the 1/4 mile time of the vehicle.
